---
title: "FINDCovTracker"
output: 
  flexdashboard::flex_dashboard:
    social: menu
    source: embed
    theme: yeti
    orientation: columns
    vertical_layout: fill
    
runtime: shiny
---

```{r setup, include=FALSE}
#Load libraries
library(data.table)
library(flexdashboard)
library(shiny)
library(dplyr)
library(lubridate)
library(stringr)
library(DT)
library(highcharter)

#Remove summarise output in console
options(dplyr.summarise.inform=F)

#Set English as default language (for lubridate date calculations)
Sys.setlocale("LC_TIME", "C")
```

```{r}
#Read dataset usig fread for large datasets
covid_dataset <- fread("https://raw.githubusercontent.com/JuanVallarta/CovTracker/main/data/data_all.csv")
covid_dataset <- covid_dataset %>%
    filter(set=="country") %>% 
  mutate(Date_format=lubridate::as_date(str_replace(time," ","-"))) %>% 
  #Rename some countries for the maps
  mutate(name=case_when(
    name=="United States"~"United States of America",
    name=="Republic of the Congo"~"Republic of Congo",
    name=="Myanmar (Burma)"~"Myanmar",
    name=="Serbia"~"Republic of Serbia",
    name=="Bosnia & Herzegovina"~"Bosnia and Herzegovina",
    name=="North Macedonia"~"Macedonia",
    TRUE~name
  ))

###FUNCTIONS###
#Function to summarize average rates and number of days per metric, period and countries#
summarise_data <- function(metric,time_period,countries,date1,date2){
  summarised_dataset <- covid_dataset %>%
    #Filter by selected countries
    filter(name %in% countries) %>% 
    #Filter by selected dates
    filter(time>=date1 & time<=date2) %>% 
    #Remove NA and 0 values
    filter(!is.na(!!sym(metric))&!!sym(metric)!=0) %>% 
    #Group observations by time period
    mutate(Date=floor_date(as_date(time),time_period)) %>% 
    group_by(Country=name,Date)%>% 
    #Summarize number of days and average rates
    summarise("Days"=n(),
              "Rates"= round(mean(!!sym(metric)),4)) %>%
    #Arrange by country and date
    arrange(Country,Date) %>% 
    #Format date variable
    mutate(Date=paste(lubridate::year(Date),str_to_title(as.character(lubridate::month(Date,label=T,abbr=F))),sep=" "))
  
  #Transform months into quartes, if selected
  if(time_period=="quarter"){
    summarised_dataset <- summarised_dataset %>%
      mutate(Date=str_replace_all(Date,c("January"="Quarter 1","April"="Quarter 2","July"="Quarter 3","October"="Quarter 4")))
  }
  #Create a new variable with a date format
  summarised_dataset <- summarised_dataset %>%
    mutate(Date_format=lubridate::ym(str_replace(Date," ","-")))

  return(summarised_dataset)
}

#Function to summarize average rates and number of days per country and time period (maps)#
summarise_data_map <- function(dataset){
  summarised_dataset_map <- dataset %>% 
    group_by(Country)%>%
    #Summarize number of days and average rates
    summarise("Days"=round(mean(Days),4),
               "Rates"= round(mean(Rates),4))
  return(summarised_dataset_map)
}

#Function to create interactive boxplots#
create_boxplot <- function(dataset,outcome,time_period){
  #Create a boxplot object (highcharter library)
  fig <- data_to_boxplot(dataset,!!sym(outcome),Country,group_var="Country",add_outliers=F)

  #Create the boxplot 
  fig_boxplot <- highchart() %>%
    #Zoom chart
    hc_chart(zoomType="xy") %>% 
    #Call the boxplot object
    hc_add_series_list(fig)%>%
    #Define x axis
    hc_xAxis(type="category",title=list(text="Countries")) %>%
    #Define y axis
    hc_yAxis(title=list(text=paste0(outcome, " per ",time_period))) %>%
    hc_add_series(data=dataset,type="scatter",hcaes(x=Country,y=!!sym(outcome),group=Country)) %>%
    #Match data points with boxplot
    hc_plotOptions(
      scatter=list(jitter=list(x=0.1,y=0)),
      series=list(grouping=F)
      ) %>% 
    #Define data points style
    hc_plotOptions(scatter=list(color="black",marker=list(radius=2,symbol="circle",lineWidth=0.5))) %>% 
    #Remove legend
    hc_legend(enabled=F)
  
  return(fig_boxplot)
}


#Function to plot time series lines
create_timeseries <- function(dataset,outcome,time_period){
  fig_timeseries <-
    #Create line chart
    hchart(dataset,"line",hcaes(x=Date_format,y=!!sym(outcome),group=Country)) %>% 
  #Make legend compare all countries at same time when hover
  hc_tooltip(crosshairs=T,shared=T,borderWidth=1) %>% 
  #Zoom chart
  hc_chart(zoomType="xy") %>% 
  #Defin x axis
  hc_xAxis(title=list(text="Dates")) %>%
  #Define y axis
  hc_yAxis(title=list(text=paste0(outcome, " per ",time_period)))
  
  return(fig_timeseries)
}

#Function to create the maps
create_map <- function(dataset,outcome,time_period){
  fig_map <- highchart() %>% 
    #Define map
    hc_add_series_map(worldgeojson,data_avg_rate_map(),name=paste0("Average ", outcome," per ", time_period),value=outcome,joinBy=c("name","Country")) %>% 
    #Color map
    hc_colorAxis(stops=color_stops()) %>% 
    #Set navigation for the map
    hc_mapNavigation(enabled=T)
}
```

Column {.sidebar data-width=300}
=====================================
Welcome to the FIND Covid-19 tracker. This interactive tool allows you to observe the daily number of Covid-19 tests, cases and deaths, and the corresponding average rates per capita (1000 people), reported by each country on a monthly and quarterly basis.
<br/>
Data is updated on a daily basis from https://github.com/finddx/FINDCov19TrackerData
<br/>

```{r}
#Set the widgets' options
#Metrics
vars_metric <- c("Tests","Cases","Deaths")
#Period
vars_period <- c("Month"="month","Quarter"="quarter")
#Countries
vars_country <- sort(unique(covid_dataset$name))
#Date ranges
vars_date_min <- min(covid_dataset$Date_format)
vars_date_max <- max(covid_dataset$Date_format)

#Set the widgets
#Metric
selectInput(inputId="var_metric",label="Select the metric",choices=vars_metric,selected="Tests")
#Period
selectInput(inputId="var_period",label="Select the period",choices=vars_period,selected="quarter")
#Countries
selectizeInput(inputId="var_countries",label="Select the countries (maximum 6)",choices=vars_country,multiple=T,options=list(maxItems=6),selected=c("Switzerland","United States of America"))
#Dates
sliderInput(inputId="var_dates",label="Date range",min=vars_date_min,max=vars_date_max,value=c(vars_date_min,vars_date_max))
```

Tables {data-icon="fa-table"} 
=====================================
Column {data-width=350}
-------------------------------------
### Number of reported days
```{r}
#Create a reactive dataset
data_daily_num <- reactive ({
  #Select variable based on metric
  if (input$var_metric=="Tests"){
    var_daily_num <- "new_tests_orig"
  } else if (input$var_metric=="Cases"){
    var_daily_num <- "new_cases_orig"
  } else if (input$var_metric=="Deaths"){
    var_daily_num <- "new_deaths_orig"
  }
  
  #Summarize the dataset
  summarise_data(var_daily_num,input$var_period,input$var_countries,input$var_dates[1],input$var_dates[2])
})

#Render table
DT::renderDataTable({
  DT::datatable(data_daily_num()[,c("Country","Date","Days")],rownames=F,options=list(pageLength=15,dom="ftp"))
})
```

Column {data-width=350}
-------------------------------------
### Averages rates per capita
```{r}
#Create a reactive dataset
data_avg_rate <- reactive ({
  #Select variable based on metric
  if (input$var_metric=="Tests"){
    var_avg_rate <- "cap_new_tests"
  } else if (input$var_metric=="Cases"){
    var_avg_rate <- "cap_new_cases"
  } else if (input$var_metric=="Deaths"){
    var_avg_rate <- "cap_new_deaths"
  }
  #Summarize the dataset
  summarise_data(var_avg_rate,input$var_period,input$var_countries,input$var_dates[1],input$var_dates[2])
})

#Render table
DT::renderDataTable({
  DT::datatable(data_avg_rate()[,c("Country","Date","Rates")],rownames=F,options=list(pageLength=15,dom="ftp"))
})
```

Plots {data-icon="fa-chart-pie"} 
=====================================
Column {data-width=350}
-------------------------------------
### Boxplot number of reported days
```{r}
#Create reactive boxplot
boxplot_daily_num <- reactive ({
  create_boxplot(data_daily_num(),"Days",input$var_period)
})

#Render boxplot
highcharter::renderHighchart({
  boxplot_daily_num() 
})
```

### Time series of reported days
```{r}
#Create reactive time series
timeseries_daily_num <- reactive ({
  create_timeseries(data_daily_num(),"Days",input$var_period)
})
highcharter::renderHighchart({
  timeseries_daily_num() 
})
```

Column {data-width=350}
-------------------------------------
### Boxplot average rates per capita
```{r}
#Create reactive boxplot
boxplot_avg_rate <- reactive ({
  create_boxplot(data_avg_rate(),"Rates",input$var_period)
})
#Render boxplot
highcharter::renderHighchart({
  boxplot_avg_rate() 

})
```

### Time series average rates per capita
```{r}
timeseries_avg_rate <- reactive ({
  create_timeseries(data_avg_rate(),"Rates",input$var_period)
})
highcharter::renderHighchart({
  timeseries_avg_rate() 
})
```

Maps {data-icon="fa-map"} 
=====================================
### Map average reported days
```{r}
#Create a reactive dataset
data_daily_num_map <- reactive ({
  #Summarize the dataset
  summarise_data_map(data_daily_num())
})
#Create reactive map
map_daily_num <- reactive ({
  create_map(data_daily_num_map(),"Days",input$var_period)
})
#Render map
highcharter::renderHighchart({
  map_daily_num()
})
```

### Map average rates per capita
```{r}
#Create a reactive dataset
data_avg_rate_map <- reactive ({
  #Summarize the dataset
  summarise_data_map(data_avg_rate())
})
#Create reactive map
map_avg_rate <- reactive ({
  create_map(data_avg_rate_map(),"Rates",input$var_period)
})
#Render map
highcharter::renderHighchart({
  map_avg_rate()
})
```

